Momentum in Round-Based First-Person-Shooter (FPS) E-Sports

### 2022-07-15 presentation.qmd
game theory q
6 official plots

---
title: "On Round-Based Games in First-Person-Shooter (FPS) E-Sports"
author: "Tony ElHabr"
format:
  revealjs:
    incremental: false
    theme: [dark, theme.scss]
execute: 
  include: false
  echo: false
---

```{r}
#| label: setup
library(dplyr)
library(tibble)
library(qs)
library(scales)

rounds <- qs::qread(file.path('cod_rounds.qs'))
```

## What?

-   Call of Duty League (CDL)
    -   Search and Destroy (SnD)
-   Valorant Pro League
    -   Plant/Defuse

## How does a team win a round of SnD?

@@

::: fragment
Often the defensive team will try to eliminate all offensive team members prior to making the defuse, but in some cases, they may try to "ninja" defuse.
:::

@@

## Data set

```{r}
#| label: descriptive-cod-snd-stats
n_series <- rounds |> 
  distinct(year, sheet, series) |> 
  nrow() |> 
  scales::comma()

@@

offensive_win_pct <- rounds |> 
  filter(is_offense) |> 
  count(win_round) |> 
  mutate(prop = n / sum(n)) |> 
  filter(win_round) |> 
  pull(prop) |> 
  scales::percent(accuracy = 1)

@@

offensive_plant_and_win_pct <- rounds |> 
  filter(is_offense, !is.na(plant)) |> 
  count(win_round) |> 
  mutate(prop = n / sum(n)) |> 
  filter(win_round) |> 
  pull(prop) |> 
  scales::percent(accuracy = 1)

offensive_first_blood_win_pct <- rounds |> 
  filter(!is.na(earned_fb)) |> 
  filter(earned_fb) |> 
  count(win_round) |> 
  mutate(prop = n / sum(n)) |> 
  filter(win_round) |> 
  pull(prop) |> 
  scales::percent(accuracy = 1)
```

@@

::: fragment
-   Offensive sides plant %: **`r plant_pct`**
:::

::: fragment
-   Offensive win %: **`r offensive_win_pct`**
:::

::: fragment
-   Win % when planting: **`r offensive_plant_and_win_pct`**
:::

::: fragment
-   Win % when earning first blood: **`r offensive_first_blood_win_pct`**
:::

## How do teams play in an information-neutral setting (round 1)?

![](offensive_win_r1_prop.png){fig-align="center"}

>>>>

---
title: "Momentum in Round-Based First-Person-Shooter (FPS) E-Sports"
author: "Tony ElHabr"
format:
  revealjs:
    incremental: false
    theme: [dark, theme.scss]
execute: 
  include: false
  echo: false
  eval: false
---

```{r}
#| label: setup
#| eval: true
library(dplyr)
library(dplyr)
library(qs)
library(scales)

rounds <- qs::qread('cod_rounds.qs')
```

```{r}
#| label: setup-eval
library(tidyverse)
library(extrafont)
library(ggtext)
library(ggsci)
blackish_background <- '#191919' ## to match slide background
gray_points <- '#4d4d4d'
gray_text <- '#999999'

font <- 'Titillium Web'
extrafont::loadfonts(quiet = TRUE)
theme_set(theme_minimal())
theme_update(
  text = element_text(family = font),
  title = element_text(size = 14, color = 'white'),
  plot.title = element_text(face = 'bold', size = 20, color = 'white'),
  plot.title.position = 'plot',
  plot.subtitle = element_text(size = 14, color = '#f1f1f1'),
  axis.text = element_text(color = 'white', size = 14),
  axis.title = element_text(size = 14, color = 'white', face = 'bold', hjust = 0.99),
  axis.line = element_blank(),
  panel.grid.major = element_line(color = gray_points),
  panel.grid.minor = element_line(color = gray_points),
  panel.grid.minor.x = element_blank(),
  panel.grid.minor.y = element_blank(),
  strip.text = element_text(color = 'white', size = 16, face = 'bold'),
  plot.margin = margin(10, 10, 10, 10),
  plot.background = element_rect(fill = blackish_background, color = blackish_background),
  plot.caption = element_text(size = 12, color = 'white', hjust = 1),
  plot.caption.position = 'plot',
  plot.tag = ggtext::element_markdown(size = 12, color = 'white', hjust = 0),
  plot.tag.position = c(0.01, 0.01),
  panel.background = element_rect(fill = blackish_background, color = blackish_background)
)
update_geom_defaults('text', list(family = font, size = 10 / .pt, fontface = 'bold'))
update_geom_defaults('point', list(color = 'white'))
update_geom_defaults('segment', list(color = 'white'))
```

## Game types

::: fragment
-   Search and Destroy (SnD) in Call of Duty (COD)
    -   2020-2022 Call of Duty League (CDL) matches (qualifiers and majors)
    -   4v4.
    -   First to 6 round wins
    -   Teams swap sides (offense and defense) after every round
:::

::: fragment
-   Plant/Defuse in Valorant
    -   Data from Valorant professional league, updated through January 1, 2022
    -   5v5
    -   First to 13 round wins
    -   One team plays offense in 12 consecutive rounds before swapping sides
    -   Must win by 2 rounds
:::

## How does a team win a round of COD SnD / Valorant plant/defuse?

@@

::: fragment
Often the defensive team will try to eliminate all o team members prior to making the defuse, but in some cases, they may try to "ninja" defuse.
:::

@@

## COD SnD data overview

```{r}
#| label: descriptive-cod-snd-stats
#| eval: true
n_series <- rounds |> 
  distinct(year, sheet, series) |> 
  nrow() |> 
  scales::comma()

@@

o_win_prop_const <- rounds |> 
  filter(is_offense) |> 
  count(win_round) |> 
  mutate(prop = n / sum(n)) |> 
  filter(win_round) |> 
  pull(prop)
o_win_pct <- scales::percent(o_win_prop_const, accuracy = 1)

@@

o_plant_and_win_pct <- rounds |> 
  filter(is_offense, !is.na(plant)) |> 
  count(win_round) |> 
  mutate(prop = n / sum(n)) |> 
  filter(win_round) |> 
  pull(prop) |> 
  scales::percent(accuracy = 1)

o_first_blood_win_pct <- rounds |> 
  filter(!is.na(earned_fb)) |> 
  filter(earned_fb) |> 
  count(win_round) |> 
  mutate(prop = n / sum(n)) |> 
  filter(win_round) |> 
  pull(prop) |> 
  scales::percent(accuracy = 1)
```

@@

::: fragment
-   Offensive plant %: **`r plant_pct`**
:::

::: fragment
-   Offensive win %: **`r o_win_pct`**
:::

::: fragment
-   Win % when planting: **`r o_plant_and_win_pct`**
:::

::: fragment
-   Win % when earning first blood: **`r o_first_blood_win_pct`**
:::

## What are your chances of winning a round, given the series state?

```{r}
#| label: cod_heatmap_prep
summarize_rounds <- function(rounds, ...) {
  rounds |> 
    group_by(pre_cumu_w, pre_cumu_l, ...) |> 
    summarize(
      n = n(),
      across(c(win_round, win_series), sum)
    ) |> 
    ungroup() |> 
    mutate(
      win_round_prop = win_round / n,
      win_series_prop = win_series / n
    )
}

n_rounds_side <- rounds |> summarize_rounds(is_offense)
o_win_prop <- n_rounds_side |> 
  filter(is_offense) |> 
  mutate(
    diff_win_round_prop = win_round_prop - !!o_win_prop_const
  )

e_o_win_series_prop <- crossing(
  pre_cumu_w = 0:5,
  pre_cumu_l = 0:5
) |> 
  mutate(
    n_remain_max = 11 - (pre_cumu_w + pre_cumu_l),
    n_remain_w = 6 - pre_cumu_w,
    ## can i account for offense in this by rotating?
    e_win_series_prop = map2_dbl(n_remain_w, n_remain_max, ~sum(dbinom(..1:..2, ..2, 0.5)))
  )

o_win_prop <- o_win_prop |> 
  inner_join(
    e_o_win_series_prop |> select(pre_cumu_w, pre_cumu_l, e_win_series_prop),
    by = c('pre_cumu_w', 'pre_cumu_l')
  ) |> 
  mutate(
    diff_win_series_prop = win_series_prop - e_win_series_prop
  )

common_heatmap_layers <- function(...) {
  list(
    ...,
    guides(
      fill = 'none'
    ),
    scale_x_continuous(
      labels = 0:5,
      breaks = seq(.5, 5.5, by = 1),
      expand = c(0, 0)
    ),
    scale_y_continuous(
      labels = 0:5,
      breaks = seq(.5, 5.5, by = 1),
      expand = c(0, 0),
      sec.axis = sec_axis(
        trans = I, 
        name = ' ', 
        breaks = seq(0.5, 5.5, by = 1), 
        labels = rep('', 6)
      )
    ),
    theme(
      panel.grid.major = element_blank(),
      axis.title = element_text(hjust = 0.5),
      plot.title = ggtext::element_markdown(hjust = 0.5),
      axis.text = element_text(size = 16, face = 'bold')
    ),
    labs(
      caption = 'CDL SnD Major and Weekly matches, 2020 - present',
      tag = '**Viz**: @TonyElHabr | **Data**: @IOUTurtle',
      x = "Offensive Team's # of Pre-Round Wins",
      y = "Defensive Team's # of Pre-Round Wins"
    )
  )
}

.dpi <- 300
save_heatmap <- function(
    p, 
    filename = paste0(str_remove(deparse(substitute(p)), '^p_'), '.png'), 
    width = 6, 
    height = 6,
    units = 'px',
    dpi = .dpi,
    ...
) {
  ggsave(
    plot = p,
    filename = filename,
    width = width * dpi,
    height = height * dpi,
    units = units,
    dpi = dpi,
    ...
  )
}

heatmap_tile <- function(..., .fill = '') {
  list(
    ...,
    geom_rect(
      color = blackish_background,
      aes(
        fill = .data[[.fill]],
        xmin = pre_cumu_w, 
        ymin = pre_cumu_l,
        xmax = pre_cumu_w + 1,
        ymax = pre_cumu_l + 1
      )
    )
  )
}

heatmap_seq_labeller <- function(.data, .label) {
  sprintf(
    '%s\n(%s/%s)', 
    scales::percent(.data[[.label]], accuracy = 1), 
    scales::comma(.data$win_round), 
    scales::comma(.data$n)
  )
}

heatmap_text_layer <- function(..., .op, .data, .label, .color, .threshold, .labeller) {
  list(
    ...,
    geom_text(
      data = .data |> filter(`.op`(.data[[.label]], .threshold)),
      # family = font,
      color = .color,
      size = 11 / .pt,
      fontface = 'bold',
      aes(
        x = pre_cumu_w + 0.5, 
        y = pre_cumu_l + 0.5,
        label = .labeller(.data = .data, .label = .label)
      )
    )
  )
}

add_text_layers <- function(..., .f) {
  list(
    ...,
    .f(.emphasize = 'low'),
    .f(.emphasize = 'high')
  )
}
```

```{r}
#| label: p_o_round_win_prop
add_o_round_win_text_layer <- function(..., .emphasize = 'low') {
  if(.emphasize == 'low') {
    op <- `<`
    .color <- blackish_background
  } else {
    op <- `>=`
    .color <- 'white'
  }
  list(
    ...,
    heatmap_text_layer(
      .data = o_win_prop,
      .op = op,
      .label = 'win_round_prop',
      .color = .color,
      .threshold = 0.52,
      .labeller = heatmap_seq_labeller
    )
  )
}

p_o_round_win_prop <- o_win_prop |> 
  ggplot() +
  common_heatmap_layers() +
  heatmap_tile(.fill = 'win_round_prop') +
  add_text_layers(.f = add_o_round_win_text_layer) +
  ggsci::scale_fill_material('teal') +
  labs(
    title = 'Offensive Round Win %'
  )
p_o_round_win_prop
save_heatmap(p_o_round_win_prop, width = 7.5, height = 5)
```

::: notes
-   This should hint at what I'm calling the "4-4" phenomenon, where the o round win % is dramatically higher than adjacent scenarios (3-4, 4-3, 5-4, 4-5, and even 3-3 or 5-5).
-   Note that round win %s are lower than one might expect in 0-5 and 1-5 states.
:::

![](o_round_win_prop.png){fig-align="center"}

## How do the empirical round results compare to expectation?

```{r}
#| label: p_o_round_win_prop_over_expected
heatmap_div_labeller <- function(.data, .label) {
  sprintf(
    '%+.0f\n(%s/%s)', 
    100 * .data[[.label]], 
    scales::comma(.data$win_round), 
    scales::comma(.data$n)
  )
}

add_o_round_win_over_expected_text_layer <- function(..., .emphasize) {
  if(.emphasize == 'low') {
    op <- `<`
    .color <- 'white'
  } else {
    op <- `>=`
    .color <- blackish_background
  }
  list(
    ...,
    heatmap_text_layer(
      .data = o_win_prop,
      .op = op,
      .label = 'diff_win_round_prop',
      .color = .color,
      .threshold = -0.1,
      .labeller = heatmap_div_labeller
    )
  )
}

extract_ggsci_mid_color <- function(palette) {
  ggsci::pal_material(palette, n = 3)(2)[2]
}

max_diff_win_round_prop <- max(abs(o_win_prop$diff_win_round_prop))
p_o_round_win_prop_over_expected <- o_win_prop |> 
  ggplot() +
  common_heatmap_layers() +
  heatmap_tile(.fill = 'diff_win_round_prop') +
  add_text_layers(.f = add_o_round_win_over_expected_text_layer) +
  scale_fill_gradient2(
    low = extract_ggsci_mid_color('purple'),
    high = extract_ggsci_mid_color('green'),
    mid = 'white',
    midpoint = 0,
    limits = c(
      -max_diff_win_round_prop, 
      +max_diff_win_round_prop
    )
  ) +
  labs(
    title = 'Offensive Round Win % Over Expected'
  )
p_o_round_win_prop_over_expected
save_heatmap(p_o_round_win_prop_over_expected, width = 7.5, height = 5)
```

::: notes
-   This should emphasize the ends of the spectrum seen on the previous graph.
:::

![](o_round_win_prop_over_expected.png){fig-align="center"}

## What are your chances of winning the series, given the series state?

```{r}
#| label: p_o_series_win_prop
add_o_series_win_text_layer <- function(..., .emphasize) {
  if(.emphasize == 'low') {
    op <- `<`
    .color <- blackish_background
  } else {
    op <- `>=`
    .color <- 'white'
  }
  list(
    ...,
    heatmap_text_layer(
      .data = o_win_prop,
      .op = op,
      .label = 'win_series_prop',
      .color = .color,
      .threshold = 0.52,
      .labeller = heatmap_seq_labeller
    )
  )
}

p_o_series_win_prop <- o_win_prop |> 
  ggplot() +
  common_heatmap_layers() +
  heatmap_tile(.fill = 'win_series_prop') +
  add_text_layers(.f = add_o_series_win_text_layer) +
  ggsci::scale_fill_material('deep-orange') +
  labs(
    title = 'Offensive Series Win %'
  )
p_o_series_win_prop
save_heatmap(p_o_series_win_prop, width = 7.5, height = 5)
```

::: notes
-   This is mostly just for reference. I think the next slide should have more insight, since it's harder to have a prior understanding about the probability of winning the series given a round state.
:::

![](o_series_win_prop.png){fig-align="center"}

## How do the empirical series results compare to expectation?

```{r}
#| label: p_o_series_win_prop_over_expected
add_o_series_win_over_expected_text_layer <- function(..., .emphasize) {
  if(.emphasize == 'low') {
    op <- `<`
    .color <- blackish_background
  } else {
    op <- `>=`
    .color <- 'white'
  }
  list(
    ...,
    heatmap_text_layer(
      .data = o_win_prop,
      .op = op,
      .label = 'diff_win_series_prop',
      .color = .color,
      .threshold = 0.08,
      .labeller = heatmap_div_labeller
    )
  )
}

max_diff_win_series_prop <- max(abs(o_win_prop$diff_win_series_prop))
p_o_series_win_prop_over_expected <- o_win_prop |> 
  ggplot() +
  common_heatmap_layers() +
  heatmap_tile(.fill = 'diff_win_series_prop') +
  add_text_layers(.f = add_o_series_win_over_expected_text_layer) +
  scale_fill_gradient2(
    low = extract_ggsci_mid_color('brown'),
    high = extract_ggsci_mid_color('cyan'),
    mid = 'white',
    midpoint = 0,
    limits = c(
      -max_diff_win_series_prop, 
      +max_diff_win_series_prop
    )
  ) +
  labs(
    title = 'Offensive Series Win % Over Expected'
  )
p_o_series_win_prop_over_expected
save_heatmap(p_o_series_win_prop_over_expected, width = 7.5, height = 5)
```

::: notes
-   Note the 4-4 and 2-4 series win percentages.
:::

![](o_series_win_prop_over_expected.png){fig-align="center"}

## Are offensive win %'s different in round 1?

```{r}
#| label: p_o_win_r1_prop
calculate_o_round_win_prop <- function(df) {
  df |> 
    count(year, game, map, is_offense, win_round) |> 
    group_by(map, is_offense) |> 
    mutate(total = sum(n), prop = n / total) |> 
    ungroup() |>
    filter(win_round, is_offense) |>
    select(
      year,
      game,
      map,
      rounds_won = n,
      total_rounds = total,
      win_round_prop = prop
    ) |>
    arrange(desc(win_round_prop))
}

devtools::source_url('https://raw.githubusercontent.com/dgrtwo/splittestr/master/R/vectorized-prop-test.R')
summarize_o_round_win_prop <- function(.round) {

  specified_round <- rounds |> 
    filter(round == .round) |> 
    calculate_o_round_win_prop()

  post_rounds <- rounds |> 
    filter(round > .round) |> 
    calculate_o_round_win_prop()

  specified_round |> 
    inner_join(
      post_rounds |> rename_with(~sprintf('post_%s', .x), -c(year, game, map)),
      by = c('year', 'game', 'map')
    ) |> 
    mutate(round = .round, .after = 'map') |> 
    arrange(desc(win_round_prop)) |> 
    mutate(
      prop_test = vectorized_prop_test(rounds_won, total_rounds, post_rounds_won, post_total_rounds),
      p_value = prop_test$p.value
    ) |> 
    arrange(p_value) |> 
    mutate(
      diff_prop = post_win_round_prop - win_round_prop,
      dir = factor(sign(diff_prop)),
      is_significant = p_value < 0.05,
      label = sprintf('%s - %s (%s)', map, game, year)
    )
}

o_win_rx_prop_pal <- c(
  `-1` = extract_ggsci_mid_color('pink'),
  `1` = extract_ggsci_mid_color('teal')
)

arw <- arrow(length = unit(5, 'pt'), type = 'closed')

plot_round_vs_post_round_slope_chart <- function(.round) {
  o_win_rx_prop <- summarize_o_round_win_prop(.round)
  x1_label <- sprintf('Round %d<br/>o Win %%', .round)
  x2_label <- sprintf('Post-Round %d<br/>o Win %%', .round)

  min_o_win_rx_prop <- o_win_rx_prop |> slice_min(diff_prop, n = 1)
  max_o_win_rx_prop <- o_win_rx_prop |> slice_max(diff_prop, n = 1)
  minmax_labels <- c(min_o_win_rx_prop$label, max_o_win_rx_prop$label)

  p <- o_win_rx_prop |> 
    ggplot() +
    geom_vline(
      data = tibble(),
      aes(xintercept = c(1, 2)), 
      color = gray_points
    ) +
    geom_hline(
      data = tibble(),
      aes(yintercept = c(0.25, 0.5, 0.75)), 
      linetype = 2,
      color = gray_points
    ) +
    geom_segment(
      data = o_win_rx_prop |> 
        mutate(
          big_segment = label %in% minmax_labels
        ),
      show.legend = FALSE,
      aes(
        x = 1,
        xend = 2,
        y = win_round_prop,
        yend = post_win_round_prop,
        color = dir,
        size = big_segment,
        group = label
      )
    ) +
    geom_text(
      data = tibble(
        y = c(0.25, 0.75)
      ) |> 
        mutate(
          lab = sprintf('%s%%', round(100 * y))
        ),
      aes(x = 2, y = y, label = lab), 
      size = 12 / .pt,
      hjust = -0.05,
      vjust = -0.25,
      fontface = 'bold',
      color = 'white'
    ) +
    ggtext::geom_richtext(
      fill = blackish_background, 
      label.color = NA,
      data = tibble(
        x = c(1, 2),
        y = rep(0.7, 2),
        lab = c(x1_label, x2_label)
      ),
      aes(x = x, y = y, label = lab), 
      size = 9 / .pt,
      hjust = 0.5,
      vjust = 0.5,
      family = font,
      fontface = 'bold',
      color = 'white'
    ) +
    ggrepel::geom_text_repel(
      data = o_win_rx_prop |> 
        filter(label %in% !!minmax_labels),
      aes(x = 1, y = win_round_prop, label = label),
      family = font,
      fontface = 'bold',
      color = 'white',
      size = 9 / .pt,
      hjust = 'right', 
      nudge_x = -0.1,
      direction = 'y'
    ) +
    ggrepel::geom_text_repel(
      data = o_win_rx_prop |> 
        filter(!(label %in% !!minmax_labels)),
      aes(x = 2, y = post_win_round_prop, label = label),
      family = font,
      color = gray_text,
      size = 9 / .pt,
      hjust = 'left', 
      nudge_x = 0.1,
      direction = 'y'
    ) +
    scale_x_continuous(
      position = 'top',
      breaks = c(1, 2),
      expand = expansion(mult = 0.5)
    ) +
    scale_y_continuous(
      limits = c(0.25, 0.75)
    ) +
    scale_size_manual(
      values = c(
        `FALSE` = 0.5,
        `TRUE` = 2
      )
    ) +
    scale_color_manual(
      name = NULL,
      values = o_win_rx_prop_pal
    ) +
    theme(
      plot.title = element_text(hjust = 0.5),
      panel.grid.major = element_blank(),
      axis.text.x = element_blank(),
      axis.text.y = element_blank()
    ) +
    labs(
      x = NULL,
      y = NULL
    )
  list(
    plot = p,
    data = o_win_rx_prop,
    largest_neg_diff_prop = min_o_win_rx_prop$diff_prop,
    largest_pos_diff_prop = max_o_win_rx_prop$diff_prop
  )
}

res_r1 <- plot_round_vs_post_round_slope_chart(1)

p_o_win_r1_vs_post_round_prop <- res_r1$plot +
  geom_curve(
    data = tibble(),
    aes(x = 1.2, y = 0.65, xend = 1.1, yend = 0.63),
    curvature = 0.2, 
    color = o_win_rx_prop_pal[["-1"]], 
    arrow = arw
  ) +
  geom_curve(
    data = tibble(),
    aes(x = 1.24, y = 0.27, xend = 1.1, yend = 0.27),
    curvature = 0-.2, 
    color = o_win_rx_prop_pal[["1"]], 
    arrow = arw
  ) +
  geom_text(
    data = tibble(
      x = c(1.2, 1.24),
      y = c(0.65, 0.27),
      hjust = c(0, 0),
      vjust = c(0.5, 0.5),
      lab = c(
        sprintf(
          'Largest negative percent difference in round 1\nvs. all other round win percentages (%+.0f%%).', 
          100 * res_r1$largest_neg_diff_prop
        ),
        sprintf(
          'Largest positive percent difference (%+.0f%%).\nOnly significant difference (p value < 0.01).',
          100 * res_r1$largest_pos_diff_prop
        )
      )
    ),
    color = 'white',
    size = 9 / .pt,
    aes(x = x, y = y, hjust = hjust, vjust = vjust, label = lab)
  )
p_o_win_r1_vs_post_round_prop

save_slope_chart <- function(
    p, 
    filename = paste0(str_remove(deparse(substitute(p)), '^p_'), '.png'), 
    width = 7.5, 
    height = 5,
    units = 'px',
    dpi = .dpi,
    ...
) {

  ggsave(
    plot = p,
    filename = filename,
    width = 7.5 * .dpi,
    height = 5 * .dpi,
    units = units,
    dpi = dpi,
    ...
  )
}

save_slope_chart(p_o_win_r1_vs_post_round_prop)
```

![](o_win_r1_vs_post_round_prop.png){fig-align="center"}

::: notes
-   This might be considered an "information-neutral" setting
:::

## Are offensive win %'s different in round 2?

```{r}
res_r2 <- plot_round_vs_post_round_slope_chart(2)
res_r2$plot
save_slope_chart(
  res_r2$plot,
  filename = 'o_win_r2_vs_post_round_prop.png'
)
```

![](o_win_r2_vs_post_round_prop.png){fig-align="center"}

::: notes
-   no trend
:::

## What are the most common series winning states?

## How do end states compare to expectation?

::: notes
-   There should be evidence of 6-0's and 6-5's being more frequent than expected given 50-50% chances. Might need to exclude major matchups, so that series are basically random.
:::

## Is there evidence for the "hot hand"?

```{r}
#| label: round_win_prop_after_b2b2b_wins
## http://keyonvafa.com/hot-hand/
get_post_streak_prob <- function(n, k, p = 0.5) {
  tosses <- rbinom(n, 1, p)
  runs <- rle(tosses)
  n_neg_after <- length(which(runs$values == 1 & runs$lengths >= k))
  n_pos_after <- sum(runs$lengths[which(runs$values == 1 & runs$lengths >= k)] - k)

  ## edge case
  if (n %in% cumsum(runs$lengths)[which(runs$values == 1 & runs$lengths >= k)]) {
    n_neg_after <- n_neg_after - 1
  }

  n_pos_after / (n_pos_after + n_neg_after)
}

simulate_post_streak_prob <- function(sims = 1000, ...) {
  rerun(
    sims,
    get_post_streak_prob(...)
  ) |> 
    flatten_dbl() |> 
    mean(na.rm = TRUE)
}
# prob_n6_k2 <- simulate_post_streak_prob(sims = 10000, n = 6, k = 2, p = 0.5)
prob_n6_k3 <- simulate_post_streak_prob(sims = 10000, n = 6, k = 3, p = 0.5)

round_streaks <- rounds |> 
  filter(round <= 6) |> 
  group_by(year, sheet, series, team) |> 
  mutate(
    won_prior_round2 = lag(win_round, n = 2, default = NA),
    won_prior_round3 = lag(win_round, n = 3, default = NA)
  ) |> 
  ungroup() |>
  select(
    year,
    sheet,
    series,
    team,
    is_offense,
    win_round,
    won_prior_round,
    won_prior_round2,
    won_prior_round3
  )

round_streaks_after_b2b <- round_streaks |> 
  filter(!is.na(won_prior_round2)) |> 
  count(win_round, won_prior_round, won_prior_round2, sort = TRUE)

round_win_prop_after_b2b_wins <- round_streaks_after_b2b |> 
  filter(won_prior_round, won_prior_round2) |> 
  mutate(
    total = sum(n),
    prop = n / total
  ) |> 
  filter(win_round) |> 
  select(n, total, prop)

round_streaks_after_b2b2b <- round_streaks |> 
  filter(!is.na(won_prior_round2), !is.na(won_prior_round3)) |> 
  count(win_round, won_prior_round, won_prior_round2, won_prior_round3, sort = TRUE)

round_win_prop_after_b2b2b_wins <- round_streaks_after_b2b2b |> 
  filter(won_prior_round, won_prior_round2, won_prior_round3) |> 
  mutate(
    total = sum(n),
    prop = n / total
  ) |> 
  filter(win_round) |> 
  select(n, total, prop)
```

```{r}
#| label: round_win_prop_after_b2b2b_wins-eval
#| eval: true
prob_n6_k3 <- 0.3838163
round_win_prop_after_b2b2b_wins <- tibble(
  n = 352,
  total = 632
) |> 
  mutate(prop = n / total)
```

::: fragment
-   Round win % after a streak of 3 round wins:
    -   Expected: **`r 100 * round(prob_n6_k3, 3)`%**
    -   Actual: **`r 100 * round(round_win_prop_after_b2b2b_wins$prop, 3)`%** (`r round_win_prop_after_b2b2b_wins$n` / `r round_win_prop_after_b2b2b_wins$total` )
:::

::: notes
-   Note this is sort of biased by the fact that teams end up winning series in 6 rounds more often than expected.
:::

## What is the optimal strategy?

::: notes
-   plant at most prevalent/winningest site? plant at other site? go for kills?
:::
